---
title: "Untitled"
author: "Thomas Chuffart"
date: "27 septembre 2018"
output: beamer_presentation
fig_width: 3 
fig_height: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## CPI and GDP: get data from FRED Database
\tiny
```{r, echo=TRUE, results='hide', message=FALSE}
  library(tidyverse)
  library(data.table)
  library(viridis)
  library(quantmod)
  my.names <- data.table(var=c("CPILFESL","GDPC1"), name=c("CPI","GDP"))
  df= getSymbols('CPILFESL',src='FRED', auto.assign=F) 
  df = data.frame(date=time(df), coredata(df))
  df.gdp_us =getSymbols('GDPC1',src='FRED', auto.assign=F) 
  df.gdp_us = data.frame(date=time(df.gdp_us), coredata(df.gdp_us))

  df<-merge(df,df.gdp_us,by="date")
  dt<-data.table(df)
  dt %>% gather(var,value,-date) %>% data.table() -> dt2
  dt2<-merge(dt2,my.names,by="var")

  # Base 100 pour la date t = 1 

  dt2=dt2[,id:=1:.N, by=var]  
  dt2=dt2[,var0:=100*value/sum(ifelse(id==1,value,0)),by=var] 
```

## CPI et GDP Figure

\tiny
```{r, echo=TRUE, fig1, out.width = '70%'}
  ggplot(data=dt2,aes(x=date,y=var0,color=name,linetype=name))+geom_line(size=1.1)+
    scale_y_log10(breaks=c(100,200,400,800))+theme_minimal()+
    theme(plot.caption=element_text(hjust=0),legend.position="top")+guides(linetype=F)+
    scale_color_viridis(name="Variable",discrete=T,end=0.8)
```

## CPI expectation change et GDP change: get data from FRED

\tiny 
```{r, echo=TRUE, results='hide', message=FALSE}
rm(list = ls())
my.names <- data.table(var=c("MICH","A191RL1Q225SBEA"),
                       name=c("Inflation_Expectation","GDP Change"))
df= getSymbols('MICH',src='FRED', auto.assign=F) 
df = data.frame(date=time(df), coredata(df))
df.gdp_us =getSymbols('A191RL1Q225SBEA',src='FRED', auto.assign=F) 
df.gdp_us = data.frame(date=time(df.gdp_us), coredata(df.gdp_us))

df<-merge(df,df.gdp_us,by="date")
dt<-data.table(df)
dt %>% gather(var,value,-date) %>% data.table() -> dt2
dt2<-merge(dt2,my.names,by="var")

```
## Michigan inflation expectation et GDP change Figure

\tiny
```{r, echo=TRUE, fig2, out.width = '70%'}
ggplot(data=dt2,aes(x=date,y=value,color=name,linetype=name))+
  geom_line(size=1.1)+
  theme_minimal()+theme(plot.caption=element_text(hjust=0),
                        legend.position="top")+
  guides(linetype=F)+
  scale_color_viridis(name="Variable",discrete=T,end=0.8)
```

## Exchange rate China/US in level and % change: get data
\tiny
```{r, echo=TRUE, results='hide', message=FALSE}
rm(list = ls())
# set dates
date.start <- "2006-01-01"
date.end <- "2015-12-31"
my.names <- data.table(var=c("DEXCHUS"),
                       name=c("Exchange rates"))
df= getSymbols('DEXCHUS',src='FRED', auto.assign=F) 
df <- df[!(is.na(df$DEXCHUS)),]
df <- df[paste(date.start,date.end,sep="/")]
df.exrate_change <- log(df$DEXCHUS) - log(lag(df$DEXCHUS))
df.exrate_change <- data.frame(date=time(df.exrate_change), coredata(df.exrate_change))

df = data.frame(date=time(df), coredata(df))
dt<-data.table(df)
dt %>% gather(var,value,-date) %>% data.table() -> dt2
dt2<-merge(dt2,my.names,by="var")

```

## Exchange rate China/US in level 
\tiny
```{r, echo=TRUE, fig3, out.width = '70%'}
  ggplot(data=dt2,aes(x=date,y=value,color=name,linetype=name))+geom_line(size=1.1)+
  theme_minimal()+theme(plot.caption=element_text(hjust=0),legend.position="top")+
  guides(linetype=F)+scale_color_viridis(name="Variable",discrete=T,end=0.8)

```

## Exchange rate China/US % change
\tiny
```{r, echo=TRUE, fig4, out.width='70%'}
  plot(df.exrate_change,  type="l")
```

## Tendance déterministe 1
\tiny
```{r, echo=TRUE, fig5, out.width = '80%'}
  T=500; 
  e = rnorm(T)
  y = rep(NA,T)
  y[1] = 0
  for (t in 1:T){
    y[t] = 0.05*t + e[t]
  }
  plot(y,type='l')
```

## Tendance déterministe 2
\tiny
```{r, echo=TRUE, fig6, out.width = '80%'}
  t <- c(1:T)
  model_deter <- lm(y~0+ t)
  plot(model_deter$residuals, type = 'l')
```

## Marche aléatoire 1
\tiny
```{r, echo=TRUE, fig7, out.width = '80%'}
  T=500; 
  e = rnorm(T)
  y = rep(NA,T)
  y[1] = 0
  for (t in 2:T){
    y[t] = e[t]+ y[t-1] 
  }
  plot(y,type='l')
```

## Marche aléatoire 2
\tiny
```{r, echo=TRUE, fig8, out.width = '80%'}
  T=500; 
  e = rnorm(T,0,10)
  y = rep(NA,T)
  y[1] = 0
  for (t in 2:T){
    y[t] = e[t]+ y[t-1] 
  }
  plot(y,type='l')
```

## Marche aléatoire 3
\tiny
```{r, echo=TRUE, fig9, out.width = '80%'}
  T=500; 
  e = rnorm(T)
  y = rep(NA,T)
  y[1] = 0
  for (t in 2:T){
    y[t] = 0.1 + e[t]+ y[t-1] 
  }
  plot(y,type='l')
```

## Marche aléatoire 4
\tiny
```{r, echo=TRUE, fig10, out.width = '80%'}
  T=500; 
  e = rnorm(T)
  y = rep(NA,T)
  y[1] = 0
  for (t in 2:T){
    y[t] = e[t]+ y[t-1] 
  }
  plot(y-lag(y),type='l')
```

## Spurious regression
\tiny
```{r, echo=TRUE}
N = 1000
T=500 
t_value = rep(NA,N)
p_value = rep(NA,N)
for (i in 1:N){
  e = rnorm(T)
  u = rnorm(T)
  y = rep(NA,T)
  x = rep(NA,T)
  y[1] = 0
  x[1] = 0
  for (t in 2:T){
    y[t] = e[t]+ y[t-1] 
    x[t] = x[t-1] + u[t]
  }
  results_sum <- summary(lm(y~x))
  results_coeff <- results_sum$coefficients
  beta.estimate <- results_coeff["x", "Estimate"] 
  std.error <- results_coeff["x", "Std. Error"]  
  t_value[i] <- beta.estimate/std.error  
  p_value[i] <- 2*pt(-abs(t_value[i]), df=T-1)  
}
sum(p_value<0.05)/N
```

## Spurious regression: figure
```{r, echo=TRUE, fig11, out.width = '80%'}
  plot(density(t_value))
  abline(v=1.96,col="blue")
  abline(v=-1.96,col="blue")
```

## Choc deterministe 
\tiny
```{r, echo=TRUE}

rm(list = ls())

T = 100
e = rnorm(T)
y = rep(NA,T)
y[1] = 0

for (t in 2:50){
  y[t] = 0.2*t + 0.8 *y[t-1] + e[t]
}
for (t in 51:T){
  y[t] = 0.2*t + 0.8 *y[t-1] 
}
```
## autre frame

```{r, echo=TRUE, fig12, out.width = '80%'}

plot(y,type='l')

```




