---
title: "Untitled"
author: "Thomas Chuffart"
date: "15 Octobre 2019"
output:
  beamer_presentation:
  fig_caption: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## CPI and GDP: get data from FRED Database
\tiny
```{r, echo=TRUE, results='hide', message=FALSE}
  library(tidyverse)
  library(data.table)
  library(viridis)
  library(tseries)
  library(quantmod)
  my.names <- data.table(var=c("CPILFESL","GDPC1"), name=c("CPI","GDP"))
  df= getSymbols('CPILFESL',src='FRED', auto.assign=F) 
  df = data.frame(date=time(df), coredata(df))
  df.gdp_us =getSymbols('GDPC1',src='FRED', auto.assign=F) 
  df.gdp_us = data.frame(date=time(df.gdp_us), coredata(df.gdp_us))

  df<-merge(df,df.gdp_us,by="date")
  dt<-data.table(df)
  dt %>% gather(var,value,-date) %>% data.table() -> dt2
  dt2<-merge(dt2,my.names,by="var")

  # Base 100 pour la date t = 1 

  dt2=dt2[,id:=1:.N, by=var]  
  dt2=dt2[,var0:=100*value/sum(ifelse(id==1,value,0)),by=var] 
```

## CPI et GDP Figure

\tiny
```{r, echo=TRUE, CPI_GDP, fig.cap='US CPI and GDP', fig.env='figure', fig.height=4.5}
  ggplot(data=dt2,aes(x=date,y=var0,color=name,linetype=name))+geom_line(size=1.1)+
    scale_y_log10(breaks=c(100,200,400,800))+theme_minimal()+
    theme(plot.caption=element_text(hjust=0),legend.position="top")+guides(linetype=F)+
    scale_color_viridis(name="Variable",discrete=T,end=0.8)
```

## CPI expectation change et GDP change: get data from FRED

\tiny 
```{r, echo=TRUE, results='hide', message=FALSE}
rm(list = ls())
my.names <- data.table(var=c("MICH","A191RL1Q225SBEA"),
                       name=c("Inflation_Expectation","GDP Change"))
df= getSymbols('MICH',src='FRED', auto.assign=F) 
df = data.frame(date=time(df), coredata(df))
df.gdp_us =getSymbols('A191RL1Q225SBEA',src='FRED', auto.assign=F) 
df.gdp_us = data.frame(date=time(df.gdp_us), coredata(df.gdp_us))

df<-merge(df,df.gdp_us,by="date")
dt<-data.table(df)
dt %>% gather(var,value,-date) %>% data.table() -> dt2
dt2<-merge(dt2,my.names,by="var")

```
## Michigan inflation expectation et GDP change Figure

\tiny
```{r, echo=TRUE, CPI_GDP_change, fig.cap='US CPI and GDP change', fig.env='figure', fig.height=4.5}
ggplot(data=dt2,aes(x=date,y=value,color=name,linetype=name))+
  geom_line(size=1.1)+
  theme_minimal()+theme(plot.caption=element_text(hjust=0),
                        legend.position="top")+
  guides(linetype=F)+
  scale_color_viridis(name="Variable",discrete=T,end=0.8)
```

## Exchange rate China/US in level and % change: get data
\tiny
```{r, echo=TRUE, results='hide', message=FALSE}
rm(list = ls())
# set dates
date.start <- "2006-01-01"
date.end <- "2015-12-31"
my.names <- data.table(var=c("DEXCHUS"),
                       name=c("Exchange rates"))
df= getSymbols('DEXCHUS',src='FRED', auto.assign=F) 
df <- df[!(is.na(df$DEXCHUS)),]
df <- df[paste(date.start,date.end,sep="/")]
df.exrate_change <- log(df$DEXCHUS) - log(lag(df$DEXCHUS))
df.exrate_change <- data.frame(date=time(df.exrate_change), coredata(df.exrate_change))

df = data.frame(date=time(df), coredata(df))
dt<-data.table(df)
dt %>% gather(var,value,-date) %>% data.table() -> dt2
dt2<-merge(dt2,my.names,by="var")

```

## Exchange rate China/US in level 
\tiny
```{r, echo=TRUE, exrate_china, fig.cap='China exchange rates', fig.env='figure', fig.height=4.5}
  ggplot(data=dt2,aes(x=date,y=value,color=name,linetype=name))+geom_line(size=1.1)+
  theme_minimal()+theme(plot.caption=element_text(hjust=0),legend.position="top")+
  guides(linetype=F)+scale_color_viridis(name="Variable",discrete=T,end=0.8)

```

## Exchange rate China/US % change
\tiny
```{r, echo=TRUE, exrate_perc, fig.cap='China exchange rates %', fig.env='figure', fig.height=4.5}
  plot(df.exrate_change,  type="l")
```

## Processus AR, mu = 0, phi1 = 0.2
\tiny

```{r, eval=TRUE, echo=TRUE, AR1_1, fig.cap="AR(1) with $\\phi_1 = 0.2$", fig.env='figure', fig.height=5}
  T <- 100
  mu <- 0
  phi1 <- 0.2
  e <- rnorm(T)
  y <- rep(NA,T)
  y[1] = 0
  for (t in 2:T){
    y[t] <- mu + phi1*y[t-1] + e[t]
  }
  plot(y,type='l',
        xlab="Time")
```


## Processus AR, mu = 0, AR1_2 = -0.2
\tiny
```{r, eval=TRUE, echo=TRUE, AR1_2, fig.cap="AR(1) with $\\phi_1 = -0.2$", fig.env='figure', fig.height=5}
  T <- 100
  mu <- 0
  phi1 <- -0.2
  e <- rnorm(T)
  y <- rep(NA,T)
  y[1] = 0
  for (t in 2:T){
    y[t] <- mu + phi1*y[t-1] + e[t]
  }
  plot(y,type='l')
```

## Processus AR, mu = 0, phi1 = 0.9
\tiny
```{r, eval=TRUE, echo=TRUE, AR1_3, fig.cap="AR(1) with $\\phi_1 = 0.9$", fig.env='figure', fig.height=5}
  T <- 100
  mu <- 0
  phi1 <- 0.9
  e <- rnorm(T)
  y <- rep(NA,T)
  y[1] = 0
  for (t in 2:T){
    y[t] <- mu + phi1*y[t-1] + e[t]
  }
  plot(y,type='l')
```


## Processus AR, mu = 0, phi1 = -0.9
\tiny
```{r, eval=TRUE, echo=TRUE, AR1_4, fig.cap="AR(1) with $\\phi_1 = -0.9$", fig.env='figure', fig.height=5}
  T <- 100
  mu <- 0
  phi1 <- -0.9
  e <- rnorm(T)
  y <- rep(NA,T)
  y[1] = 0
  for (t in 2:T){
    y[t] <- mu + phi1*y[t-1] + e[t]
  }
  plot(y,type='l')
```


## Processus AR, mu = 0, phi1 = 1.01
\tiny
```{r, eval=TRUE, echo=TRUE, AR1_5, fig.cap="AR(1) with $\\phi_1 = 1.01$", fig.env='figure', fig.height=4.5}
  T <- 100
  mu <- 0
  phi1 <- 1.01
  e <- rnorm(T)
  y <- rep(NA,T)
  y[1] = 0
  for (t in 2:T){
    y[t] <- mu + phi1*y[t-1] + e[t]
  }
  plot(y,type='l')
```

## ACF AR(1)
\tiny
```{r, echo=FALSE, ACF_AR1, fig.cap='ACF AR(1)', fig.env='figure', fig.height=4.5}
  T <- 100
  mu <- 0
  phi11 <- 0.2
  phi12 <- 0.9
  phi13 <- -0.2
  phi14 <- -0.9
  e <- rnorm(T)
  y1 <- rep(NA,T)
  y1[1] = 0
  y2 <- rep(NA,T)
  y2[1] = 0  
  y3 <- rep(NA,T)
  y3[1] = 0  
  y4 <- rep(NA,T)
  y4[1] = 0  
  for (t in 2:T){
    y1[t] <- mu + phi11*y1[t-1] + e[t]
    y2[t] <- mu + phi12*y2[t-1] + e[t]
    y3[t] <- mu + phi13*y3[t-1] + e[t]
    y4[t] <- mu + phi14*y4[t-1] + e[t]
  }
  par(mar=c(1,1,1,1))  
  par(mfrow=c(2,2))
  acf(y1)
  acf(y2)
  acf(y3)
  acf(y4)
```

## PACF AR(1)
\tiny
```{r, echo=FALSE, PACF_AR1, fig.cap='PACF AR(1)', fig.env='figure', fig.height=4.5}
  T <- 100
  mu <- 0
  phi11 <- 0.2
  phi12 <- 0.9
  phi13 <- -0.2
  phi14 <- -0.9
  e <- rnorm(T)
  y1 <- rep(NA,T)
  y1[1] = 0
  y2 <- rep(NA,T)
  y2[1] = 0  
  y3 <- rep(NA,T)
  y3[1] = 0  
  y4 <- rep(NA,T)
  y4[1] = 0  
  for (t in 2:T){
    y1[t] <- mu + phi11*y1[t-1] + e[t]
    y2[t] <- mu + phi12*y2[t-1] + e[t]
    y3[t] <- mu + phi13*y3[t-1] + e[t]
    y4[t] <- mu + phi14*y4[t-1] + e[t]
  }
  par(mar=c(1,1,1,1))  
  par(mfrow=c(2,2))
  pacf(y1)
  pacf(y2)
  pacf(y3)
  pacf(y4)
```

## PACF AR(2)
\tiny
```{r, echo=FALSE, PACF_AR2, fig.cap='ACF et PACF AR(2)', fig.env='figure', fig.height=4.5}
  T <- 100
  mu <- 0
  phi11 <- 0.2
  phi21 <- 0.6
  phi12 <- -0.2
  phi22 <- 0.6
  e <- rnorm(T)
  y1 <- rep(NA,T)
  y1[1] = 0
  y1[2] = 0
  y2 <- rep(NA,T)
  y2[1] = 0  
  y2[2] = 0  
  for (t in 3:T){
    y1[t] <- mu + phi11*y1[t-1] + phi21*y1[t-2] + e[t]
    y2[t] <- mu + phi12*y2[t-1] + phi22*y2[t-2] + e[t]
  }
  par(mar=c(1,1,1,1))  
  par(mfrow=c(2,2))  
  acf(y1)
  acf(y2)
  pacf(y1)
  pacf(y2)
```

## ACF MA(1)
\tiny
```{r, echo=FALSE, MA_plot, fig.cap='MA(1), $\\theta_{11} = 0.2$, $\\theta_{12} = 0.9$, $\\theta_{13} = -0.2$, $\\theta_{14} = -0.9$', fig.env='figure', fig.height=4.5}
  T <- 100
  mu <- 0
  theta11 <- 0.2
  theta12 <- 0.9
  theta13 <- -0.2
  theta14 <- -0.9
  e <- rnorm(T)
  y1 <- rep(NA,T)
  y1[1] = 0
  y2 <- rep(NA,T)
  y2[1] = 0  
  y3 <- rep(NA,T)
  y3[1] = 0  
  y4 <- rep(NA,T)
  y4[1] = 0  
  for (t in 2:T){
    y1[t] <- mu + theta11*e[t-1] + e[t]
    y2[t] <- mu + theta12*e[t-1] + e[t]
    y3[t] <- mu + theta13*e[t-1] + e[t]
    y4[t] <- mu + theta14*e[t-1] + e[t]
  }
  par(mar=c(1,1,1,1))  
  par(mfrow=c(2,2))
  plot(y1, type = "l")
  plot(y2, type = "l")
  plot(y3, type = "l")
  plot(y4, type = "l")

  
```

## ACF MA(1)
\tiny
```{r, echo=FALSE, MA_acf, fig.cap='ACF MA(1)', fig.env='figure', fig.height=4.5}
  T <- 100
  mu <- 0
  theta11 <- 0.2
  theta12 <- 0.9
  theta13 <- -0.2
  theta14 <- -0.9
  e <- rnorm(T)
  y1 <- rep(NA,T)
  y1[1] = 0
  y2 <- rep(NA,T)
  y2[1] = 0  
  y3 <- rep(NA,T)
  y3[1] = 0  
  y4 <- rep(NA,T)
  y4[1] = 0  
  for (t in 2:T){
    y1[t] <- mu + theta11*e[t-1] + e[t]
    y2[t] <- mu + theta12*e[t-1] + e[t]
    y3[t] <- mu + theta13*e[t-1] + e[t]
    y4[t] <- mu + theta14*e[t-1] + e[t]
  }
  par(mar=c(1,1,1,1))  
  par(mfrow=c(2,2))
  acf(y1)
  acf(y2)
  acf(y3)
  acf(y4)

  
```

## PACF MA(1)
\tiny
```{r, echo=FALSE, MA_pacf, fig.cap='PACF MA(1)', fig.env='figure', fig.height=4.5}
  T <- 100
  mu <- 0
  theta11 <- 0.2
  theta12 <- 0.9
  theta13 <- -0.2
  theta14 <- -0.9
  e <- rnorm(T)
  y1 <- rep(NA,T)
  y1[1] = 0
  y2 <- rep(NA,T)
  y2[1] = 0  
  y3 <- rep(NA,T)
  y3[1] = 0  
  y4 <- rep(NA,T)
  y4[1] = 0  
  for (t in 2:T){
    y1[t] <- mu + theta11*e[t-1] + e[t]
    y2[t] <- mu + theta12*e[t-1] + e[t]
    y3[t] <- mu + theta13*e[t-1] + e[t]
    y4[t] <- mu + theta14*e[t-1] + e[t]
  }
  par(mar=c(1,1,1,1))  
  par(mfrow=c(2,2))
  pacf(y1)
  pacf(y2)
  pacf(y3)
  pacf(y4)

  
```

## Tendance déterministe 1
\tiny
```{r, echo=TRUE, t_deter, fig.cap='Tendance deterministe 1', fig.env='figure', fig.height=4.5}
  T=500
  e = rnorm(T)
  y = rep(NA,T)
  y[1] = 0
  for (t in 1:T){
    y[t] = 0.05*t + e[t]
  }
  plot(y,type='l')
```

## Tendance déterministe 2
\tiny
```{r, echo=TRUE, t_deter2, fig.cap='Tendance deterministe 2', fig.env='figure', fig.height=4.5}
  t <- c(1:T)
  model_deter <- lm(y~0+ t)
  plot(model_deter$residuals, type = 'l')
```

## Marche aléatoire 1
\tiny
```{r, echo=TRUE, random_walk_1, fig.cap='Random Walk 1', fig.env='figure', fig.height=4.5}
  T=500; 
  e = rnorm(T)
  y = rep(NA,T)
  y[1] = 0
  for (t in 2:T){
    y[t] = e[t]+ y[t-1] 
  }
  plot(y,type='l')
```

## Marche aléatoire 2
\tiny
```{r, echo=TRUE, random_walk_2, fig.cap='Random Walk 2', fig.env='figure', fig.height=4.5}
  T=500; 
  e = rnorm(T,0,10)
  y = rep(NA,T)
  y[1] = 0
  for (t in 2:T){
    y[t] = e[t]+ y[t-1] 
  }
  plot(y,type='l')
```

## Marche aléatoire 3
\tiny
```{r, echo=TRUE, random_walk_3, fig.cap='Random Walk 3', fig.env='figure', fig.height=4.5}
  T=500; 
  e = rnorm(T)
  y = rep(NA,T)
  y[1] = 0
  for (t in 2:T){
    y[t] = 0.1 + e[t]+ y[t-1] 
  }
  plot(y,type='l')
```

## Marche aléatoire 4
\tiny
```{r, echo=TRUE, random_walk_4, fig.cap='Random Walk 4', fig.env='figure', fig.height=4.5}
  T=500; 
  e = rnorm(T)
  y = rep(NA,T)
  y[1] = 0
  for (t in 2:T){
    y[t] = e[t]+ y[t-1] 
  }
  plot(y-lag(y),type='l')
```

## Spurious regression
\tiny
```{r, echo=TRUE}
N = 1000
T=500 
t_value = rep(NA,N)
p_value = rep(NA,N)
for (i in 1:N){
  e = rnorm(T)
  u = rnorm(T)
  y = rep(NA,T)
  x = rep(NA,T)
  y[1] = 0
  x[1] = 0
  for (t in 2:T){
    y[t] = e[t]+ y[t-1] 
    x[t] = x[t-1] + u[t]
  }
  results_sum <- summary(lm(y~x))
  results_coeff <- results_sum$coefficients
  beta.estimate <- results_coeff["x", "Estimate"] 
  std.error <- results_coeff["x", "Std. Error"]  
  t_value[i] <- beta.estimate/std.error  
  p_value[i] <- 2*pt(-abs(t_value[i]), df=T-1)  
}
sum(p_value<0.05)/N
```

## Spurious regression: figure
```{r, echo=TRUE, spurious, fig.cap='Spurious regression', fig.env='figure', fig.height=4.5}
  plot(density(t_value))
  abline(v=1.96,col="blue")
  abline(v=-1.96,col="blue")
```

## Choc deterministe 
\tiny
```{r, echo=TRUE}

rm(list = ls())

T = 100
e = rnorm(T)
y = rep(NA,T)
y[1] = 0

for (t in 2:50){
  y[t] = 0.2*t + 0.8 *y[t-1] + e[t]
}
for (t in 51:T){
  y[t] = 0.2*t + 0.8 *y[t-1] 
}
```

## Choc deterministe 
```{r, echo=TRUE, choc, fig.cap='Choc déterministe', fig.env='figure', fig.height=4.5}

plot(y,type='l')

```




